The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl

### 2018-04-25 results.tex
apr 25
jpcl draft

\section{Results and discussion}\label{sec:rnd}

\subsection{Normal Auger decay}\label{ssec:na}

The KL$_{2,3}$L$_{2,3}$ normal Auger decay following K-shell ionization of aqueous \ki~and \cli~can be written as follows
%
\begin{align*}
\text{h}\nu + \text{K}^{+}_{\text{aq}} \rightarrow 
	\text{K}^{2+}_{\text{aq}} (1s^{-1}) + e^{-}_{\text{ph}}\rightarrow
	\text{K}^{3+}_{\text{aq}} (2p^{-2}) + e^{-}_{\text{ph}} + e^{-}_{\text{Auger}}\\
\text{h}\nu + \text{Cl}^{-}_{\text{aq}} \rightarrow
	\text{Cl}^{0}_{\text{aq}} (1s^{-1}) + e^{-}_{\text{ph}}\rightarrow
	\text{Cl}^{+}_{\text{aq}}(2p^{-2}) + e^{-}_{\text{ph}} + e^{-}_{\text{Auger}}
\end{align*}
%
It populates the 2p$^{-2}$($^3$P, $^1$D, $^1$S) final states. The $^3$P final states have a very low intensity since the corresponding transitions are forbidden from angular momentum and parity conservation rules. %However, these states are not observed in the experimental spectra presented on Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl} because they overlap with the 2p$^{-2}$($^1$D) and 2p$^{-2}$($^1$D)$V'$ states in the kinetic energy region where they are expected to occur. 
In the case of K$^{+}_{\text{aq}}$ the maxima of the $^1$S and $^1$D KL$_{2,3}$L$_{2,3}$ Auger lines at photon energy h$\nu = 3616.0$\,eV are located at 2958.0\,eV and 2968.5\,eV kinetic energy, respectively (see Fig.\ \ref{fg:2dmap_k}). For Cl$^{-}_{\text{aq}}$, the lines corresponding to the Cl$^{+}$ 2p$^{-2}$($^1$S) and ($^1$D) states are located at 2373.1\,eV and 2382.0\,eV kinetic energy for photon energy h$\nu = 2830.0$\,eV (see Fig.\ \ref{fg:2dmap_cl}).

@@

The KL$_{2,3}$L$_{2,3}$ normal Auger lines may disperse with photon energy close to threshold due to energy exchange between the photoelectron and Auger electron called post-collision interaction (PCI). As a result of this interaction, first, an Auger peak becomes asymmetric with a high kinetic energy tail, and second, its intensity maximum is shifted to higher kinetic energies close to threshold \citep{russek86:911,guillemin15:012503}. These effects are seen in our spectra -- the asymmetric tail of the peaks resulting from PCI is clearly shown on the Auger spectra at photon energies 3616\,eV in the case of \ki, and 2830\,eV in the case of \cli.

@@

Further, we compare the positions of the normal KLL Auger lines of both Cl$^{-}_{\text{aq}}$ and K$^{+}_{\text{aq}}$ close to threshold with those recorded far from threshold, at photon energies h$\nu = 5$\,keV (see Ref.\ \citep{ceolin17:263003} for details). In the latter case, the maxima of the $^1$D and $^1$S states appear at 2381.0\,eV and 2372.0\,eV for Cl$^{-}_{\text{aq}}$, and 2967.4\,eV and 2957.0\,eV for K$^{+}_{\text{aq}}$, respectively. The maxima were found to be shifted by $\sim$1\,eV towards lower kinetic energies as compared to the spectra measured close to threshold. The magnitude of the shift is constant in the photon energy range of $\sim$8\,eV above threshold (see Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl}) and similar for the two ions. %Moreover, it is the same for both ions suggesting that it does not depend on the initial charge of the ion. 
A possible explanation of the shift observed in our experiment is given in Ref.\ \cite{tchaplyguine07:124314}, which focuses on the Auger decay of large Kr clusters close to the 3d$_{5/2}$ ionization threshold. The observed 4p$^{-2}$($^1$S, $^1$D, $^3$P) Auger peaks were found to be shifted by 0.7\,eV to higher kinetic energies compared to the positions of the peaks far above threshold. Moreover, the shift did not vary with the photon energy close to threshold. Consequently, it was proposed that these features originate from a process of internal ionization, i.e.\ excitation of the photoelectron into the conduction band, followed by normal Auger decay. The observed shift was explained as resulting from the PCI-like interaction between the electron excited in the conduction band and the Auger electron. Further investigations are, however, planned in the case of liquid samples.

@@

\subsection{Resonant Auger decay} \label{ssec:ra}

@@

The KL$_{2,3}$L$_{2,3}$ Auger decay following resonant K-shell excitation of solvated \ki~and \cli~can be written as follows
%
\begin{align*}
\text{h}\nu + \text{K}^{+}_{\text{aq}} \rightarrow \text{K}^{+*}_{\text{aq}} (1s^{-1}V) \rightarrow \text{K}^{2+}_{\text{aq}} (2p^{-2}V') + e^{-}_{\text{Auger}}\\
%
\text{h}\nu + \text{Cl}^{-}_{\text{aq}} \rightarrow \text{Cl}^{-*}_{\text{aq}} (1s^{-1}V) \rightarrow \text{Cl}^{0}_{\text{aq}}(2p^{-2}V') + e^{-}_{\text{Auger}}
\end{align*}
%
where $V$ and $V'$ denote the unoccupied orbitals in the presence of the 1s$^{-1}$ and 2p$^{-2}$ core holes, respectively. The character of these states is discussed below.

@@

In order to rationalize the pre-edge region of the experimental XAS spectra and the differences in the AES spectra of \ki$_{\text{aq}}$ and \cli$_{\text{aq}}$, we computed the lowest core excited states of the bare \ki~and \cli~ions and their hexa-coordinated clusters. The theoretical XAS spectra are presented in Fig.\ \ref{fg:xas_kcl}. In the bare ions (lowermost panels of Fig.\ \ref{fg:xas_kcl}), the lowest energy peak corresponds to the dipole allowed 1s$^{-1}$4p state. The next dipole allowed state, 1s$^{-1}$5p, is located 4.3\,eV and 10.8\,eV higher in the cases of \ki~and \cli, respectively. Together with the two dipole allowed transitions, we show the dipole forbidden 1s$^{-1}$3d states of the bare ions as blue crosses at photon energies 3611.2\,eV in the case of \ki~and 2831.7\,eV in the case of \cli, respectively. It is noteworthy that the positions of the 1s$^{-1}$4p and 1s$^{-1}$3d states are inverted in \ki~and \cli. In the case of Cl$^{-}$ the 1s$^{-1}$4p excitation has lower energy and the 1s$^{-1}$3d excitation is close to the 1s$^{-1}$5p state. On the contrary, in K$^{+}$ the 1s$^{-1}$3d excitation has lower energy and lies below the 1s$^{-1}$4p state. We note in passing that the intensity of the Cl$^{-}$(1s$^{-1}$4p) state is lower than that of the \cli(1s$^{-1}$5p) state contrary to what is observed in \ki. This difference can be explained by the lower electron density of the 4p compared to the 5p electron in the region close to the core hole which thus results in the lower oscillator strength of the 1s$^{-1}$4p compared to the 1s$^{-1}$5p transition in \cli~(see Fig.\ \ref{fg:si_rdens_ions} in SI).

>>>>

%\section{Results and discussion}\label{sec:rnd}

%\subsection{Normal Auger decay}\label{ssec:na}

The KL$_{2,3}$L$_{2,3}$ normal Auger decay following K-shell ionization of aqueous \ki~and \cli~are shown in Fig.\ \ref{fg:auger} %can be written as follows
%
%\begin{align*}
%\text{h}\nu + \text{K}^{+}_{\text{aq}} \rightarrow 
%	\text{K}^{2+}_{\text{aq}} (1s^{-1}) + e^{-}_{\text{ph}}\rightarrow
%	\text{K}^{3+}_{\text{aq}} (2p^{-2}) + e^{-}_{\text{ph}} + e^{-}_{\text{Auger}}\\
%\text{h}\nu + \text{Cl}^{-}_{\text{aq}} \rightarrow
%	\text{Cl}^{0}_{\text{aq}} (1s^{-1}) + e^{-}_{\text{ph}}\rightarrow
%	\text{Cl}^{+}_{\text{aq}}(2p^{-2}) + e^{-}_{\text{ph}} + e^{-}_{\text{Auger}}
%\end{align*}
%
It populates the 2p$^{-2}$($^3$P, $^1$D, $^1$S) final states. The $^3$P final states have a very low intensity since the corresponding transitions are forbidden from angular momentum and parity conservation rules. %However, these states are not observed in the experimental spectra presented on Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl} because they overlap with the 2p$^{-2}$($^1$D) and 2p$^{-2}$($^1$D)$V'$ states in the kinetic energy region where they are expected to occur. 
In the case of K$^{+}_{\text{aq}}$ the maxima of the $^1$S and $^1$D KL$_{2,3}$L$_{2,3}$ Auger lines at photon energy h$\nu = 3616.0$\,eV are located at 2958.0\,eV and 2968.5\,eV kinetic energy, respectively (see Fig.\ \ref{fg:2dmap_k}). For Cl$^{-}_{\text{aq}}$, the lines corresponding to the Cl$^{+}$ 2p$^{-2}$($^1$S) and ($^1$D) states are located at 2373.1\,eV and 2382.0\,eV kinetic energy for photon energy h$\nu = 2830.0$\,eV (see Fig.\ \ref{fg:2dmap_cl}).

@@

The KL$_{2,3}$L$_{2,3}$ normal Auger lines may be shifted to higher kinetic energies and also disperse with photon energy close to threshold due to energy exchange between the photoelectron and Auger electron called post-collision interaction (PCI). %As a result of this interaction, first, an Auger peak becomes asymmetric with a high kinetic energy tail, and second, its intensity maximum is shifted to higher kinetic energies close to threshold
\citep{russek86:911,guillemin15:012503}. These effects are seen in our spectra -- the asymmetric tail of the peaks resulting from PCI is clearly shown on the Auger spectra at photon energies 3616\,eV in the case of \ki, and 2830\,eV in the case of \cli.

@@

Further, we compare the positions of the normal KLL Auger lines of both Cl$^{-}_{\text{aq}}$ and K$^{+}_{\text{aq}}$ close to threshold with those recorded far from threshold, at photon energies h$\nu = 5$\,keV \citep{ceolin17:263003}. 
%In the latter case, the maxima of the $^1$D and $^1$S states appear at 2381.0\,eV and 2372.0\,eV for Cl$^{-}_{\text{aq}}$, and 2967.4\,eV and 2957.0\,eV for K$^{+}_{\text{aq}}$, respectively. The maxima were found to be shifted by $\sim$1\,eV towards lower kinetic energies 
We observe a shift of $\sim$1\,eV of the maxima towards lower kinetic energies as compared to the spectra reported in \citep{ceolin17:263003}. The magnitude of the shift is constant in the photon energy range of $\sim$8\,eV above threshold (see Figs.\ \ref{fg:2dmap_k} and \ref{fg:2dmap_cl}) and similar for the two ions. %Moreover, it is the same for both ions suggesting that it does not depend on the initial charge of the ion. 
A possible explanation of the shift observed in our experiment is given in Ref.\ \cite{tchaplyguine07:124314} where it was proposed that a process of internal ionization, i.e.\ excitation of the photoelectron into the conduction band, followed by normal Auger decay. The observed shift was explained as resulting from the PCI-like interaction between the electron excited in the conduction band and the Auger electron.
%, which focuses on the Auger decay of large Kr clusters close to the 3d$_{5/2}$ ionization threshold. The observed 4p$^{-2}$($^1$S, $^1$D, $^3$P) Auger peaks were found to be shifted by 0.7\,eV to higher kinetic energies compared to the positions of the peaks far above threshold. Moreover, the shift did not vary with the photon energy close to threshold. Consequently, it was proposed that these features originate from a process of internal ionization, i.e.\ excitation of the photoelectron into the conduction band, followed by normal Auger decay. The observed shift was explained as resulting from the PCI-like interaction between the electron excited in the conduction band and the Auger electron. Further investigations are, however, planned in the case of liquid samples.

@@

%\subsection{Resonant Auger decay} \label{ssec:ra}

@@

The KL$_{2,3}$L$_{2,3}$ Auger decay following resonant K-shell excitation of solvated \ki~and \cli~is presented on Fig.\ \ref{fg:auger}.
%can be written as follows
%%
%\begin{align*}
%\text{h}\nu + \text{K}^{+}_{\text{aq}} \rightarrow \text{K}^{+*}_{\text{aq}} (1s^{-1}V) \rightarrow \text{K}^{2+}_{\text{aq}} (2p^{-2}V') + e^{-}_{\text{Auger}}\\
%%
%\text{h}\nu + \text{Cl}^{-}_{\text{aq}} \rightarrow \text{Cl}^{-*}_{\text{aq}} (1s^{-1}V) \rightarrow \text{Cl}^{0}_{\text{aq}}(2p^{-2}V') + e^{-}_{\text{Auger}}
%\end{align*}
%%
%where $V$ and $V'$ denote the unoccupied orbitals in the presence of the 1s$^{-1}$ and 2p$^{-2}$ core holes, respectively. The character of these states is discussed below.

@@

In order to rationalize the pre-edge region of the experimental XAS spectra and the differences in the AES spectra of \ki$_{\text{aq}}$ and \cli$_{\text{aq}}$, we computed the lowest core excited states of the bare \ki~and \cli~ions and their hexa-coordinated clusters. The theoretical XAS spectra are presented in Fig.\ \ref{fg:xas_kcl}. In the bare ions (lowermost panels of Fig.\ \ref{fg:xas_kcl}), the lowest energy peak corresponds to the dipole allowed 1s$^{-1}$4p state. The next dipole allowed state, 1s$^{-1}$5p, is located 4.3\,eV and 10.8\,eV higher in the cases of \ki~and \cli, respectively. Together with the two dipole allowed transitions, we show the dipole forbidden 1s$^{-1}$3d states of the bare ions as blue crosses at photon energies 3611.2\,eV in the case of \ki~and 2831.7\,eV in the case of \cli, respectively. It is noteworthy that the positions of the 1s$^{-1}$4p and 1s$^{-1}$3d states are inverted in \ki~and \cli. In the case of Cl$^{-}$ the 1s$^{-1}$4p excitation has lower energy and the 1s$^{-1}$3d excitation is close to the 1s$^{-1}$5p state. On the contrary, in K$^{+}$ the 1s$^{-1}$3d excitation has lower energy and lies below the 1s$^{-1}$4p state. %We note in passing that the intensity of the Cl$^{-}$(1s$^{-1}$4p) state is lower than that of the \cli(1s$^{-1}$5p) state contrary to what is observed in \ki. This difference can be explained by the lower electron density of the 4p compared to the 5p electron in the region close to the core hole which thus results in the lower oscillator strength of the 1s$^{-1}$4p compared to the 1s$^{-1}$5p transition in \cli~(see Fig.\ \ref{fg:si_rdens_ions} in SI).

### 2018-04-25 suppinfo - Copy.tex
apr 25
jpcl draft

\usepackage{color}
\usepackage{xr}
\externaldocument{suppinfo}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.8]{figures/rad_dens_kcl.eps}
\caption{Radial density distributions of the singly-occupied natural orbital occupied by the excited electron corresponding to the 1s$^{-1}$4p, 1s$^{-1}$3d and 1s$^{-1}$5p core excitations in \ki~(lower panel) and \cli~(upper panel). The insets show the region of distances relevant for the overlap with the 1s core orbital whose radial density is shown as a grey shaded area.}
\label{fg:si_rdens_ions}
\end{figure}
{\color{red}
In what follows we give a tentative explanation of the difference in the radial density distributions of the 1s$^{-1}$4p and 1s$^{-1}$5p states in \ki~and \cli. In the case of \ki, the excited electron mainly sees a $2/r$ potential. In addition, it sees a short range potential coming from the pointlike nucleus and the screening electrons. The latter one normally leads to a quantum defect different from 0. Beside this one has Rydberg series with an infinite number of states. However, in case of Cl- the outer electron does not see a Coulomb potential and the short-range potential becomes dominant. As a result we see the unusual behaviour, like e.g. a finite number of bound states (here obviously 4p). In contrast to this 3d and 5p are in the continuum.}

\begin{figure}
\centering
\includegraphics[scale=0.9]{figures/partial_cross_sec_ct_time.pdf}
\caption{Partial cross sections and charge transfer time extracted from Fig.\ \ref{fg:2dmap_cl}. The blue and red curves are obtained by integrating the area of the 2p$^{-2}$ and 2p$^{-2}$4p final states ($^1$D state region only) at each photon energy step. From these curves we determine the charge transfer time $\tau_{\text{CT}}$ according to the formula $\tau_{\text{CT}} = \tau l/d$, with $\tau$ being the Cl 1s core-hole lifetime and $l/d$ being the intensity ratio of the localized (2p$^{-2}$ 4p) and delocalized (2p$^{-2}$) states at a given excitation energy \cite{foehlisch05:373}.  The green line defines the Cl$^{-}_{\text{aq}}$(1s) ionization potential.}
\label{fg:si_ct_time}
\end{figure}

>>>>

\usepackage{color}
\usepackage{xr}
\externaldocument{master}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{The all-seeing eye of resonant Auger electron spectroscopy: a study on aqueous KCl}
%\title{Tracking the hazy boundary between resonant and normal Auger decay in near the K-edges of aqueous \ki~and \cli}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.8]{figures/rad_dens_kcl.eps}
\caption{Radial density distributions of the singly-occupied natural orbital occupied by the excited electron corresponding to the 1s$^{-1}$4p, 1s$^{-1}$3d and 1s$^{-1}$5p core excitations in \ki~(lower panel) and \cli~(upper panel). The insets show the region of distances relevant for the overlap with the 1s core orbital whose radial density is shown as a grey shaded area.}
\label{fg:si_rdens_ions}
\end{figure}

In what follows we give a simple explanation of the difference in the radial density distributions of the 1s$^{-1}$4p and 1s$^{-1}$5p states in \ki~and \cli. In the case of \ki, the excited electron mainly sees a $2/r$ potential. In addition, it sees a short range potential originating from the point-like nucleus and the screening electrons. The influence of the latter can be described by a quantum defect $\delta \ne 0$, which is almost constant for the entire infinite Rydberg series. However, in case of \cli~the outer electron does not experience a Coulomb potential and the short range potential becomes dominant. As a result of the absence of the Coulomb potential we see a different behavior in the properties of the states, like e.g.\ only a finite number of bound states (here obviously 4p) \citep{buckman94:539}. In contrast to this, the 3d and 5p states are not bound. % in the continuum.}

\begin{figure}
\centering
\includegraphics[scale=0.9]{figures/partial_cross_sec_ct_time.pdf}
\caption{Partial cross sections and charge transfer time extracted from Fig.\ \ref{fg:2dmap_cl}. The blue and red curves are obtained by integrating the area of the 2p$^{-2}$ and 2p$^{-2}$4p final states ($^1$D state region only) at each photon energy step. From these curves we determine the charge transfer time $\tau_{\text{CT}}$ according to the formula $\tau_{\text{CT}} = \tau_{c} l/d$, with $\tau_{c}$ being the Cl 1s core-hole lifetime and $l/d$ being the intensity ratio of the localized (2p$^{-2}$4p) and delocalized (2p$^{-2}$) states at a given excitation energy \cite{foehlisch05:373}.  The green line defines the Cl$^{-}_{\text{aq}}$(1s) ionization potential.}
\label{fg:si_ct_time}
\end{figure}
